---
title: "Exploration of the epigraphic formula `Dis manibus` on Latin inscriptions"
author: 
- Petra Hermankova^[Aarhus University, Denmark, https://orcid.org/0000-0002-6349-0540]
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged
---

```{r}
#install.packages(c("reticulate", "stringdist", "ape", "igraph", "tidyverse", "reticulate", "jsonlite", "leaflet"))    

library(igraph)
library(ape)
library(stringdist)
library(tidyverse)
library(reticulate)
library(jsonlite)
library(leaflet)
#use_python("/usr/local/bin/python/") # use this if you have local python installed
#use_condaenv("pip")

Sys.which("python") # checks which python do I use
```
Sources: # http://amunategui.github.io/stringdist/

```{python}
# installation of packages is recommended to do via the terminal (see in your console below), pip install name-of-the-required-package

# https://vincent.doba.fr/posts/20210407_install-fiona-on-windows/ & https://geopandas.org/getting_started/install.html Issues with geopandas on windows

#to install requirements type and run in the terminal:  pip install -r requirements.txt

import fiona
import geopandas as gpd 
import requests
import seaborn as sns
import matplotlib.colors as mcolors
import matplotlib.pyplot as plt
from scipy.stats import trapz
import scipy
#import nltk
import json
import tempun
#import sddk
import numpy as np
import pandas as pd
pd.options.display.max_columns = 1000  # to see all columns
```

# Setup

**Loading data from Sciencedata.dk**

```{r}
#list_json <- jsonlite::fromJSON("https://sciencedata.dk/public/b6b6afdb969d378b70929e86e58ad975/EDH_text_cleaned_2021-01-21.json")
#EDH <- as_tibble(list_json)
```

**Loading data locally**
```{r}
# this will work only if you have the dataset available locally
EDH <- jsonlite::fromJSON("../data/EDH_text_cleaned_2021-01-21.json")
#dir.create("../figures")
```

## Text preparation

Converting the text of inscriptions to lowercase
```{r}
EDH <- EDH %>% 
  mutate(clean_text_interpretive_word_lowercase = str_to_lower(EDH$clean_text_interpretive_word)) %>% 
  mutate(clean_text_conservative_lowercase = str_to_lower(EDH$clean_text_conservative))
```

## Preparation coordinates for mapping
```{r}
# preparation of coordinates
EDH<- EDH %>% 
  separate(col = coordinates, into = c("longitude", "latitude"), sep = ",")

EDH$latitude <- as.numeric(str_replace(EDH$latitude, pattern = "\\)", replacement=""))
EDH$longitude <- as.numeric(str_replace(EDH$longitude, pattern = "c\\(", replacement=""))
```


# Dis Manibus epigraphic formula in space and time

Typically occurs on funerary inscriptions, invocation of the underworld gods.

Canonic Versions:

D M - only in the conservative version of the text
D M S -  only in the conservative version 
Dis Manibus
Diis Manibus
Dis Manibus Sacrum
Diis Manibus Sacrum

## Extracting inscriptions with for all *typical* variants of the formula

```{r}
dis_manibus_regex_minimal <- "di{1,2}s manibus sacrum|di{1,2}s manibus"
dis_manibus_regex_maximal <- "\\bd m s\\b|\\bd m\\b|di{1,2}s manibus sacrum|di{1,2}s manibus|\\bdi{1,2} man[e|i]s\\b"

dis_manibus_int<- str_subset(EDH$clean_text_interpretive_word_lowercase, dis_manibus_regex_minimal)

dis_manibus_cons<- str_subset(EDH$clean_text_conservative_lowercase, dis_manibus_regex_maximal)


# separating them in a new attribute
EDH<- EDH %>% 
  mutate(form_dis_manibus_ins = str_extract(EDH$clean_text_interpretive_word_lowercase, dis_manibus_regex_minimal)) %>% 
  mutate(form_dis_manibus_cons = str_extract(EDH$clean_text_conservative_lowercase, dis_manibus_regex_maximal)) 

EDH$form_dis_manibus <- ifelse(grepl("\\w", EDH$form_dis_manibus_ins, ignore.case = T),  EDH$form_dis_manibus_ins, EDH$form_dis_manibus_cons)

EDH$form_dis_manibus <- ifelse(grepl("d m s", EDH$form_dis_manibus, ignore.case = T),  "dis manibus sacrum", 
                               ifelse(grepl("d m", EDH$form_dis_manibus, ignore.case = T),  "dis manibus",
                               EDH$form_dis_manibus))

table(EDH$form_dis_manibus)  

# overview
table(str_extract(EDH$clean_text_interpretive_word_lowercase, dis_manibus_regex_minimal))

table(str_extract(EDH$clean_text_conservative_lowercase, dis_manibus_regex_maximal))
```

```{r}
# capturing context - 5 words before the formula and 5 words after the formula
#EDH<- EDH %>% 
#  mutate(dis_manibus_in_context_10w = str_extract(EDH$clean_text_interpretive_word_lowercase, "(\\w*\\s){0,6}di{1,2}s manibus sacrum|di{1,2}s manibus(\\w*\\s){0,6}")) 
```

**Subsetting dataset to contain only records with formula**
```{r}
#non-empty dis manibus attribute
EDH_DM <- EDH %>% 
  filter(!is.na(form_dis_manibus))
```

**Save the dataset locally**
```{r}
EDH_DMjson <- toJSON(EDH_DM, simplifyVector = TRUE)
write(EDH_DMjson, "../data/EDH_DM.json")
```

# Exploration of Dis Manibus inscriptions

## Formulae variants

### Ratio of inscriptions with DM formula

```{r}
nrow(EDH %>% 
  filter(!is.na(form_dis_manibus)))/(nrow(EDH)/100)
```

### Type of all inscription with DM formula
```{r}
EDH_DM %>% 
  count(type_of_inscription_clean, sort = T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))
```

### Ratio of funerary inscriptions with DM formula

```{r}
epitaph_all<- length(EDH$type_of_inscription_clean[EDH$type_of_inscription_clean == "epitaph"])
epitaph_dm<- length(EDH_DM$type_of_inscription_clean[EDH_DM$type_of_inscription_clean == "epitaph"])

epitaph_dm/(epitaph_all/100)
```

### Type of the formula overall
```{r}
form_dm<- EDH_DM %>% 
  #filter(type_of_inscription_clean %in% c("epitaph", "NULL")) %>% 
  filter(!is.na(form_dis_manibus)) %>% 
  count(form_dis_manibus, sort=T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))

form_dm

write_csv(form_dm, "../output/DM_form_total_numbers.csv")
```

## Province

### Type of formula by province - frequence
```{r, fig.height=10, fig.width=14}
EDM_DM_freq<- EDH_DM %>% 
  #count(province_label_clean, form_dis_manibus, sort=T) %>% 
  ggplot(aes(y=province_label_clean)) + 
  geom_bar(aes(fill=form_dis_manibus)) +
  #coord_cartesian(xlim=c(0,1700)) +
  theme_minimal() +
  theme(text = element_text(size=12)) +
  labs(y="Roman province", x="Formula", title= "Frequence of Dis Manibus formula by province", subtitle = ggtitle(paste("n =", nrow(EDH_DM), "inscriptions (EDH)")))  
ggsave("../figures/EDM_DM_province_freq.png")
EDM_DM_freq
```

### Ratio of DM variants per province
```{r, fig.height=10}

DM_form<- EDH_DM %>% 
  select(form_dis_manibus, province_label_clean) %>% 
  count(form_dis_manibus, province_label_clean, sort=T)


EDH_DM %>% 
  ggplot(aes(y=province_label_clean, fill=form_dis_manibus)) +
  geom_bar(position="fill") +
  theme_minimal() +
  theme(text = element_text(size=12)) +
  geom_vline(xintercept=c(0.25, 0.5, 0.75), linetype="dotted", color = "blue", size=0.8) + 
  labs(y="Roman province", x="Formula", title= "Ratio of Dis Manibus formulae by province", subtitle = ggtitle(paste("n =", nrow(EDH_DM), "inscriptions (EDH)"))) 
  #geom_label(aes(label= form_dis_manibus)) +
  #geom_label(aes(label = form_dis_manibus), colour = "black", fontface = "bold", hjust = -0.1)
```

### Provinces with Dis manibus sacrum
```{r, fig.height=7}

EDH_DM$province_label_clean <-factor(EDH_DM$province_label_clean)

EDH_DM$form_dis_manibus <-factor(EDH_DM$form_dis_manibus)


EDH_DM %>% 
  count(form_dis_manibus, province_label_clean, sort=T) %>% 
  group_by(form_dis_manibus) %>%  
  filter(form_dis_manibus == "dis manibus sacrum") %>% 
  ggplot(aes(y=fct_reorder(province_label_clean, n), x=n)) +
  geom_col(fill= "blue") +
  labs(y="Roman province", x="n", title= "Frequence of 'Dis Manibus Sacrum' formula by province", subtitle = ggtitle(paste("n =", nrow(EDH_DM), "inscriptions (EDH)"))) +
  geom_text(aes(label = n), colour = "black", size= 3, hjust = -0.1)

```

```{r, fig.height=7}
EDH_DM %>% 
  count(form_dis_manibus, province_label_clean, sort=T) %>% 
  group_by(form_dis_manibus) %>%  
  filter(form_dis_manibus == "dis manibus") %>% 
  ggplot(aes(y=fct_reorder(province_label_clean, n), x=n)) +
  geom_col(fill= "blue") +
  labs(y="Roman province", x="n", title= "Frequence of 'Dis Manibus' formula by province", subtitle = ggtitle(paste("n =", nrow(EDH_DM), "inscriptions (EDH)"))) +
  geom_text(aes(label = n), colour = "black", size= 3, hjust = -0.1)
```

### Ratio of formulae per province normalized by the number of total inscriptions
```{r}

DM_form_province<- EDH_DM %>% 
  count(form_dis_manibus, province_label_clean, sort=T) %>% 
  pivot_wider(names_from = form_dis_manibus, values_from = n)

province_total_insc<- EDH %>% 
  count(province_label_clean)

provinces_DM<- merge(x=province_total_insc, y=DM_form_province, by="province_label_clean", all=T)

provinces_DM<- provinces_DM %>% 
  rename(all_insc=n)

provinces_DM <- provinces_DM %>% 
  mutate(dm = provinces_DM$`dis manibus`+provinces_DM$`diis manibus`) %>% 
  mutate(dms = provinces_DM$`dis manibus sacrum`+provinces_DM$`diis manibus sacrum`) %>% 
  mutate(dm_ratio = round(dm/(all_insc/100),2)) %>% 
  mutate(dms_ratio = round(dms/(all_insc/100),2))

provinces_DM
write_csv(provinces_DM, "../output/DM_form_provinces.csv")
```

### How common the dm and the dms formulae are (of all inscriptions from a given province)

```{r, fig.height=7}
#Filter those provinces where the ratio is higher than mean for all provinces fro a given formula

higher_dm<- provinces_DM %>% 
    filter(dm_ratio > mean(na.omit(provinces_DM$dm_ratio)))
higher_dm

```
```{r}
higher_dms<- provinces_DM %>% 
    filter(dms_ratio > mean(na.omit(provinces_DM$dms_ratio)))
higher_dms
```

```{r}
higher_dm_dms<- provinces_DM %>% 
    filter(dm_ratio > mean(na.omit(provinces_DM$dm_ratio)& dms_ratio > mean(na.omit(provinces_DM$dms_ratio))))
higher_dm_dms
```

## Mapping

### Map of Dis Manibus occurences - cummulative
```{r}
# preparation of subsets
# dis manibus formula
EDH_dis_manibus <- EDH_DM %>% 
  filter(form_dis_manibus == "dis manibus")

# diis manibus formula
EDH_diis_manibus <- EDH_DM %>% 
  filter(form_dis_manibus == "diis manibus")

# dis manibus sacrum formula
EDH_dis_manibus_sacrum <- EDH_DM %>% 
  filter(form_dis_manibus == "dis manibus sacrum")

# diis manibus sacrum formula
EDH_diis_manibus_sacrum <- EDH_DM %>% 
  filter(form_dis_manibus == "diis manibus sacrum")

```


```{r}
DM_map<- #head(EDH_DM, 100) %>% 
  leaflet(EDH_DM) %>% 
  leaflet(width="100%") %>%
  #addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
  addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
  #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
  #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  setView( lng = 12.9239625, lat = 41.9515694, zoom = 3.5 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
  #addPolylines(data = roads, color = "purple", weight = 1, opacity = 0.7) %>% 
  
  addCircles(lng = EDH_dis_manibus$longitude, 
             lat = EDH_dis_manibus$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "blue" , fillColor = "blue",) %>%
  addCircles(lng = EDH_dis_manibus_sacrum$longitude, 
             lat = EDH_dis_manibus_sacrum$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "red" , fillColor = "red",) %>%
    addCircles(lng = EDH_diis_manibus$longitude, 
             lat = EDH_diis_manibus$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "black" , fillColor = "black",) %>%
  addCircles(lng = EDH_diis_manibus_sacrum$longitude, 
             lat = EDH_diis_manibus_sacrum$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "orange" , fillColor = "orange",) %>%
  

  #addAwesomeMarkers(~EDH$longitude, ~EDH$latitude, icon=icons) %>% 
  addLegend(position = "bottomright",
  colors = c("Blue", "Green", "Red", "Orange"),
  labels = c("dis manibus", "diis manibus", "dis manibus sacrum", "diis manibus sacrum"), opacity = 1,
  title = "Dis Manibus formula (EDH)" 
) %>% 
  addScaleBar(position="bottomleft")

DM_map
```

## Demographic preferences

**Preparation of the dataset - unique people records**

```{r}
EDH_small <- EDH_DM %>% 
  select(id, people, form_dis_manibus, type_of_inscription_clean, type_of_monument_clean)

EDH_smu <- EDH_small %>% unnest_wider(people)
names(EDH_smu)

# this works only on Linux computer, not on Windows

EDH_people<- EDH_small %>% unnest_wider(people) %>% 
  unnest_longer(names(EDH_smu))
```

### Gender of person on inscription
```{r}
EDH_people<- EDH_people %>% 
  #filter(form_dis_manibus == "dis manibus" | form_dis_manibus == "dis manibus sacrum") %>% 
  mutate(gender_clean = str_replace_all(gender, "W\\?", "female")) %>%
  mutate(gender_clean = str_replace_all(gender_clean, "M\\?", "male"))


EDH_people %>% 
  count(gender_clean, form_dis_manibus, sort=T) %>% 
  mutate(freq = n/(sum(n)/100)) %>% 
  ggplot(aes(y=gender_clean, x=n)) +
  geom_col(aes(fill=form_dis_manibus)) +
  #geom_label(aes(label = round(freq, 2)), hjust=0.3, size = 3.5) +
  theme_linedraw(base_size = 12) +
  labs(y="Gender", x="Formula", title= "Dis Manibus formula by gender", subtitle = ggtitle(paste("n =", length(EDH_DM$people), "inscriptions (EDH)"))) 


```

```{r}
EDH_people %>% 
  ggplot(aes(x=gender_clean, fill=form_dis_manibus)) +
  geom_bar(position="fill") +
  #geom_label(aes(label = round(freq, 2)), hjust=0.3, size = 3.5) +
  #theme_linedraw(base_size = 12) +
  labs(y="Gender", x="Formula", title= "Ratio of Dis Manibus formula by gender", subtitle = ggtitle(paste("n =", length(EDH_DM$people), "inscriptions (EDH)")))
```


### Status of person on inscription
```{r, fig.width=14, fig.height=10}

EDH_people %>% 
  #filter(form_dis_manibus == "dis manibus" | form_dis_manibus == "dis manibus sacrum") %>% 
  mutate(status_clean = str_replace_all(status, "\\?", "")) %>% 
  count(status_clean, form_dis_manibus, sort=T) %>% 
  ggplot(aes(y=status_clean, x=n)) +
  geom_col(aes(fill=form_dis_manibus)) +
  #coord_cartesian(xlim = c(0, 15800)) +
  labs(y="Status", x="Individuals", title= "Dis Manibus formula by personal status", subtitle = ggtitle(paste("n =", nrow(EDH_people), "individuals (EDH)"))) +
  theme(legend.position = c(0.9, 0.15)) +
  scale_fill_discrete(name="Formula")
```

```{r, fig.width=14, fig.height=10}

EDH_people<- EDH_people %>% 
  #filter(form_dis_manibus == "dis manibus" | form_dis_manibus == "dis manibus sacrum") %>% 
  mutate(status_clean = str_replace_all(status, "\\?", "")) 

```

Status of a person was split into new column `status_highest` that records only the highest status category in the following order: 
1) slaves,
2) freedmen / freedwomen, 
3) Augustales,
4) military personnel, 
5) lower local offices, administration of imperial estates, 
6) decurial order, high local offices,
7) equestrian order,
8) senatorial order, 
9) rulers (foreign)

```{r}
EDH_people$status_highest <- ifelse(grepl("decurial order, higher local offices; military personnel", EDH_people$status_clean, ignore.case = T),  "decurial order, higher local offices",
                                    ifelse(grepl("equestrian order; decurial order, higher local offices", EDH_people$status_clean, ignore.case = T), "equestrian order",
                                           ifelse(grepl("senatorial order; equestrian order", EDH_people$status_clean, ignore.case = T), "senatorial order", 
                                                  ifelse(grepl("Augustales; freedmen / freedwomen", EDH_people$status_clean, ignore.case = T), "Augustales", 
                                                         ifelse(grepl("decurial order, higher local offices; Augustales", EDH_people$status_clean, ignore.case = T), "decurial order, higher local offices", 
                                                                ifelse(grepl("decurial order, higher local offices; freedmen / freedwomen", EDH_people$status_clean, ignore.case = T), "decurial order, higher local offices",
                                                                       ifelse(grepl("equestrian order; lower local offices, administration of imperial estates", EDH_people$status_clean, ignore.case = T), "equestrian order",
                                                                          
                                                  EDH_people$status_clean)))))))
```
     
                                    
```{r, fig.width=10, fig.height=5}

EDH_status_count <- EDH_status %>% 
  count(form_dis_manibus, status_clean, sort = T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2)) 

mean_dms <- mean(EDH_status_count$ratio[EDH_status_count$form_dis_manibus == "dis manibus sacrum"])

EDH_people %>% 
  ggplot(aes(y=status_highest, fill=form_dis_manibus)) +
  geom_bar(position="fill") +
  #coord_cartesian(xlim = c(0, 15800)) +
  labs(y="Status", x="Individuals", title= "Ratio of Dis Manibus formula by personal status", subtitle = ggtitle(paste("n =", nrow(EDH_people), "individuals (EDH)"))) +
  #theme(legend.position = c(0.9, 0.15)) +
  scale_fill_discrete(name="Formula") +
  geom_vline(xintercept =mean_dms/10, linetype = "longdash", size=0.3)

```

### Age of person on inscription
```{r}
EDH_people %>% 
  mutate(age_yr = as.numeric(`age: years`)) %>% 
  filter(!is.na(age_yr)) %>% 
  count(age_yr, form_dis_manibus, sort=T) %>% 
  ggplot(aes(age_yr, n, colour= form_dis_manibus)) +
  geom_point(aes(colour=form_dis_manibus)) +
  labs(y="Individuals", x="Years at death", title= "Dis Manibus formula by age in years of the deceased", subtitle = ggtitle(paste("n =", nrow(EDH_people), "individuals (EDH)")))+
  geom_smooth(se = FALSE, method = lm) +
  facet_wrap(~form_dis_manibus)
  
```


#### Average age per formula type

@continue here

```{r}
average_age <- EDH_people %>% 
  mutate(age_yr = as.numeric(`age: years`)) %>% 
  filter(!is.na(age_yr)) %>%
  select(age_yr) 

average_age_all<- mean(na.omit(average_age$age_yr))

average_age_per_formula <- EDH_people %>% 
  mutate(age_yr = as.numeric(`age: years`)) %>% 
  filter(!is.na(age_yr)) %>%
  group_by(form_dis_manibus) %>% 
  summarise_at(vars(age_yr), list(name=mean)) %>% 
  mutate(diff_av_age  =  name-average_age_all)

average_age_per_formula

```

**Saving DM with people data locally:**
```{r}
EDH_DMpeoplejson <- toJSON(EDH_people, simplifyVector = TRUE)
write(EDH_DMpeoplejson, "../data/EDH_DM_people.json")
```



### Inscribed monument
```{r}
EDH_DM %>% 
  count(type_of_monument_clean, form_dis_manibus, sort = T) %>% 
  ggplot(aes(y=type_of_monument_clean, x=n)) + 
  geom_col(aes(fill=form_dis_manibus)) +
  theme_minimal() +
  labs(y="Inscribed object", x="Inscriptions", title= "Dis Manibus formula by inscribed object", subtitle = ggtitle(paste("n =", nrow(EDH_DM), "inscriptions (EDH)")))
```

### Inscribed material
```{r}
EDH_DM %>% 
  count(material_clean, form_dis_manibus, sort = T) %>% 
  ggplot(aes(y=material_clean, x=n)) + 
  geom_col(aes(fill=form_dis_manibus)) +
  theme_minimal() +
  labs(y="Inscribed material", x="Inscriptions", title= "Dis Manibus formula by inscribed material", subtitle = ggtitle(paste("n =", nrow(EDH_DM), "inscriptions (EDH)")))
```

