title: "`Dis manibus`in time"
author: 
- Petra Hermankova^[Aarhus University, Denmark, https://orcid.org/0000-0002-6349-0540]
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged
---

```{r, echo=FALSE}
#install.packages(c("reticulate", "stringdist", "ape", "igraph", "tidyverse", "reticulate", "jsonlite", "leaflet"))    

library(igraph)
library(ape)
library(stringdist)
library(tidyverse)
library(reticulate)
library(jsonlite)
library(leaflet)
#use_python("/usr/local/bin/python/") # use this if you have local python installed
#use_condaenv("pip")

Sys.which("python") # checks which python do I use
```
Sources: # http://amunategui.github.io/stringdist/

```{python}
# installation of packages is recommended to do via the terminal (see in your console below), pip install name-of-the-required-package

# https://vincent.doba.fr/posts/20210407_install-fiona-on-windows/ & https://geopandas.org/getting_started/install.html Issues with geopandas on windows

#to install requirements type and run in the terminal:  pip install -r requirements.txt

import fiona
import geopandas as gpd 
import requests
import seaborn as sns
import matplotlib.colors as mcolors
import matplotlib.pyplot as plt
from scipy.stats import trapz
import scipy
#import nltk
import json
import tempun
#import sddk
import numpy as np
import pandas as pd
pd.options.display.max_columns = 1000  # to see all columns
```

# Setup

**Loading data from Sciencedata.dk**
```{r}
list_json <- jsonlite::fromJSON("https://sciencedata.dk/public/b6b6afdb969d378b70929e86e58ad975/formulae/EDH_DM.json")
EDH_DM <- as_tibble(list_json)
```

**Loading data locally**
```{r}
# this will work only if you have the dataset available locally
#EDH_DM <- jsonlite::fromJSON("../data/EDH_DM.json")
#dir.create("../figures")
```

# Temporal usage of the formulae

```{r}
# preparation of the dataset
EDH_DM$not_before <- as.numeric(as.character(EDH_DM$not_before))
EDH_DM$not_after <- as.numeric(as.character(EDH_DM$not_after))
```

```{python}
# How many inscriptions have both dates?
EDH_dated = r.EDH_DM[(r.EDH_DM["not_before"].notnull()) | (r.EDH_DM["not_after"].notnull())]
len(EDH_dated)
```

```{python}
# Generate a list of 1000 random dates for each inscription in the dataset
EDH_dated["random_dates"] = EDH_dated.apply(lambda row: tempun.model_date(
    row["not_before"], row["not_after"], size=1000,), axis=1)
```

```{r}
# saving as r-object
py$EDH_dated -> EDH_dated
```

```{python}
# Count random dates in temporal timeblocks in a specified time range (from 200 BC to 600 AD, temporal blocks by 25 years)
EDH_complete_simulations = tempun.timeblocks_from_randoms(
    EDH_dated, "random_dates", [-200, 600, 25])
```

## DM inscriptions in time
```{python}
# Figure 1
# plot all the dataset to see the epigraphic production in time
plt.style.use("seaborn-white")
plt.rcParams['figure.dpi'] = 300
plt.rcParams['font.size'] = 10
tab_colors_list = list(mcolors.TABLEAU_COLORS.keys())
fig, ax = plt.subplots(figsize=(14, 6), tight_layout=True)
tempun.plot_timeblocks_data(
    EDH_complete_simulations, ax=ax, color="gray", label=f"EDH (n={len(EDH_dated)})")
    
ax.set_xlabel("Year", weight="bold")
ax.set_ylabel("Count", weight="bold")
ax.set_title("Dis Manibus: Epigraphic production over time (EDH)")
ax.legend(title="Data source", title_fontsize="large", bbox_to_anchor=(1, 1), loc='upper right')
#plt.axvline(x=212, linewidth=0.5, color = 'orange', linestyle='dashed')
#fig.suptitle(f'Comparison of epigraphic production over time', fontsize=16,fontweight="bold")
fig
plt.savefig('../figures/01_EDH_Epi_production_time.png')
```


```{python}
# Inscriptions by their type over time, EDH
simulations_by_type_len_EDH = []
for ins_type in r.EDH_DM["form_dis_manibus"].unique():
    if ins_type != "NULL":
        subset_df = EDH_dated[EDH_dated["form_dis_manibus"] == ins_type]
        simulations_by_type_len_EDH.append((ins_type, len(subset_df)))
simulations_by_type_len_EDH = sorted(
    simulations_by_type_len_EDH, key=lambda x: x[1], reverse=True)
simulations_by_type_len_EDH
simulations_by_type_EDH = []
for ins_type_tup in simulations_by_type_len_EDH[:8]:
    subset_df = EDH_dated[EDH_dated["form_dis_manibus"]
                          == ins_type_tup[0]]
    simulations = tempun.timeblocks_from_randoms(
        subset_df, "random_dates", [-200, 600, 25])
    ins_type_N = len(subset_df)
    simulations_by_type_EDH.append([ins_type_tup[0], ins_type_N, simulations])
simulations_by_type_EDH = sorted(
    simulations_by_type_EDH, key=lambda x: x[1], reverse=True)
date_vars_by_instypes = []
for ins_type_tup in simulations_by_type_len_EDH[:10]:
    subset_df = EDH_dated[EDH_dated["form_dis_manibus"]
                          == ins_type_tup[0]]
    date_vars = []
    for n in range(100):
        date_vars.append(
            [date_var[n] for date_var in subset_df["random_dates"] if date_var != None])
    date_vars_by_instypes.append(date_vars)
```

```{python}
simulations_by_type_len_EDH
```

```{python}
# Figure 2
plt.style.use("seaborn-white")
plt.rcParams['figure.dpi'] = 300
plt.rcParams['font.size'] = 10
fig, ax = plt.subplots(figsize=(14, 6), tight_layout=True)
for ins_type_sims, color in zip(simulations_by_type_EDH, tab_colors_list[:len(simulations_by_type_EDH)]):
    tempun.plot_timeblocks_data(ins_type_sims[2], ax=ax, color=color)
  
    
ax.set_xlabel("Year", weight="bold")
ax.set_ylabel("Count", weight="bold")
ax.set_title(f"'Dis manibus' formula on inscriptions over time (EDH, n={len(EDH_dated)})", weight="bold")
markers = [plt.Line2D([0, 0], [0, 0], color=color, lw=4)
           for color in tab_colors_list[:len(simulations_by_type_EDH)]]
legend_labels_EDH = [tup[0] + " (n={})".format(str(tup[1])) for tup in simulations_by_type_EDH]
ax.legend(markers, legend_labels_EDH, numpoints=1,  
           title=f"Formula types (EDH n={len(EDH_dated)})", title_fontsize="large", bbox_to_anchor=(1, 1), loc='upper right')
plt.savefig('../figures/02_EDH_Typologies_comparison_time.png')
fig
```



## DM inscriptions in periods 
```{python}
# periods definitions
periods = {  # to avoid overlaps, startdates are postponed by one year, when needed
    "2/2 1st c. BCE": {"startdate": -50, "enddate": 0, "duration": 50},
    "1/2 1st c. CE": {"startdate": 1, "enddate": 50,"duration": 50},
    "2/2 1st c. CE": {"startdate": -51, "enddate": 100, "duration": 50},
    "1/2 2nd c. CE": {"startdate": 101, "enddate": 150, "duration": 50},
    "2/2 2nd c. CE": {"startdate": 151, "enddate": 200, "duration": 50},
    "1/2 3rd c. CE": {"startdate": 201, "enddate": 250,"duration": 50},
    "2/2 3rd c. CE": {"startdate": 251, "enddate": 300,"duration": 50},
    "1/2 4th c. CE": {"startdate": 301, "enddate": 350,"duration": 50}
}
timeblocks_periods = [(periods[period]["startdate"],
                       periods[period]["enddate"],
                       periods[period]["duration"]) for period in periods]
timeblocks_periods
```

```{python}
def date_to_str(date):
    if date < 0:
        date = str(abs(date)) + " BC"
    else:
        date = "AD " + str(date)
    return date
periods_labels = []
for period in periods.keys():
    start = date_to_str(periods[period]["startdate"])
    end = date_to_str(periods[period]["enddate"])
    periods_labels.append(period + "\n({0}-{1})".format(start, end))
periods_labels
```

```{python}
# loading shapefile from Pleaides for the largest extent of the Roman Empire, AD 117
# source: https://raw.githubusercontent.com/pelagios/magis-pleiades-regions/main/pleiades-regions-magis-pelagios.geojson
pleiades_regions = gpd.read_file('../data/pleiades_regions.geojson', driver='GeoJSON')
RE_merged = pleiades_regions.unary_union
```

```{python}
def get_date_var(randoms):
    try:
        return randoms[0]
    except:
        return None
# selecting one random date out of the 1000 version
EDH_dated["date_var_1"] = EDH_dated["random_dates"].apply(get_date_var)
```

```{r}
# updating the values in existing r-object
py$EDH_dated -> EDH_dated
```

```{python}
# modifying coordinates in EDH to fit the script
# EDH_dated[['Longitude', 'Latitude']] = pd.DataFrame(EDH_dated.coordinates. tolist(), index=EDH_dated.index)
```

```{python}
EDH_dated_df_by_periods = []
for period in timeblocks_periods:
    EDH_dated_sample = EDH_dated[EDH_dated["date_var_1"].between(
        period[0], period[1])]
    # tranforming EDH as geodataframe
    EDH_dated_sample = EDH_dated_sample[EDH_dated_sample["latitude"].notnull()]
    EDH_dated_sample = gpd.GeoDataFrame(EDH_dated_sample, geometry=gpd.points_from_xy(EDH_dated_sample["longitude"], EDH_dated_sample["latitude"]))
    # selecting only those dated and within borders of the Empire (the largest extent in AD 117)
    EDH_dated_sample = EDH_dated_sample[EDH_dated_sample.within(RE_merged)]
    EDH_dated_df_by_periods.append(EDH_dated_sample)
```

```{python}
# Figure 3
# plot EDH inscriptions with location in 8 plots grouped by period
plt.style.use("seaborn-white")
plt.rcParams['figure.dpi'] = 300
plt.rcParams['font.size'] = 10
fig, axs  = plt.subplots(4, 2, figsize=(7, 7), tight_layout=True)
contexts_pct = {}
    
for df_edh, ax, period in zip(EDH_dated_df_by_periods, axs.ravel(), periods_labels):
    pleiades_regions.plot(ax=ax, color="lightgray")
    df_edh.plot(markersize=0.04, color="darkblue", ax=ax, alpha=0.2, label=len(df_edh))
    ax.set_title(period, fontsize=6)
    ax.set_axis_off()
    markers = [plt.Line2D([0,0],[0,0],color=color, marker=".", linestyle="") for color in ["darkblue", "red"]]
    legend_labels = ["EDH (n={0})".format(str(len(df_edh)))]
    
    ax.legend(markers, legend_labels, numpoints=1, bbox_to_anchor=(0.6, 1), loc='upper left', fontsize=5)
   
plt.tight_layout(pad=0)
plt.subplots_adjust(wspace=0.0, hspace=0.0)
fig.suptitle(f'Spatial extent of the epigraphic production by historic period', fontsize=8, fontweight="bold")
plt.savefig('../figures/03_EDH_Formulae_production_periods.png')
fig
```

## DM variants in time - needs fixing
```{python}
dis_manibus = EDH_dated[EDH_dated["form_dis_manibus"] == "dis manibus"]
len(dis_manibus)
```

```{python}
diis_manibus = EDH_dated[EDH_dated["form_dis_manibus"] == "diis manibus"]
len(diis_manibus)
```

```{python}
dis_manibus_sac = EDH_dated[EDH_dated["form_dis_manibus"] == "dis manibus sacrum"]
len(dis_manibus_sac)
```
```{python}
diis_manibus_sac = EDH_dated[EDH_dated["form_dis_manibus"] == "diis manibus sacrum"]
len(diis_manibus_sac)
```


```{python}
EDH_dated_df_by_periods_dm = []
for period in timeblocks_periods:
    EDH_dated_sample_dm = dis_manibus[dis_manibus["date_var_1"].between(
        period[0], period[1])]
    # tranforming EDH as geodataframe
    EDH_dated_sample_dm = EDH_dated_sample_dm[EDH_dated_sample_dm["latitude"].notnull()]
    EDH_dated_sample_dm = gpd.GeoDataFrame(EDH_dated_sample_dm, geometry=gpd.points_from_xy(EDH_dated_sample_dm["longitude"], EDH_dated_sample_dm["latitude"]))
    # selecting only those dated and within borders of the Empire (the largest extent in AD 117)
    EDH_dated_sample_dm = EDH_dated_sample_dm[EDH_dated_sample_dm.within(RE_merged)]
    EDH_dated_df_by_periods_dm.append(EDH_dated_sample_dm)
```

```{python}
EDH_dated_df_by_periods_dm
```


```{python}
EDH_dated_df_by_periods_dms = []
for period in timeblocks_periods:
    EDH_dated_sample_dms = dis_manibus_sac[dis_manibus_sac["date_var_1"].between(
        period[0], period[1])]
    # tranforming EDH as geodataframe
    EDH_dated_sample_dms = EDH_dated_sample_dms[EDH_dated_sample_dms["latitude"].notnull()]
    EDH_dated_sample_dms = gpd.GeoDataFrame(EDH_dated_sample_dms, geometry=gpd.points_from_xy(EDH_dated_sample_dms["longitude"], EDH_dated_sample_dms["latitude"]))
    # selecting only those dated and within borders of the Empire (the largest extent in AD 117)
    EDH_dated_sample_dms = EDH_dated_sample_dms[EDH_dated_sample_dms.within(RE_merged)]
    EDH_dated_df_by_periods_dms.append(EDH_dated_sample_dms)
```

```{python}
EDH_dated_df_by_periods_dms
```


```{python}
# Figure 4
# plot EDH inscriptions with location in 8 plots grouped by period
plt.style.use("seaborn-white")
plt.rcParams['figure.dpi'] = 300
plt.rcParams['font.size'] = 10
fig, axs  = plt.subplots(4, 2, figsize=(7, 7), tight_layout=True)
contexts_pct = {}
    
for df_edh_dms, df_edh_dm, ax, period in zip(EDH_dated_df_by_periods_dms, EDH_dated_df_by_periods_dm, axs.ravel(), periods_labels):
    pleiades_regions.plot(ax=ax, color="lightgray")
    df_edh_dm.plot(markersize=0.04, color="darkblue", ax=ax, alpha=0.5, label=len(df_edh_dms))
    df_edh_dms.plot(markersize=0.04, color="red", ax=ax, alpha=0.5, label=len(df_edh_dm))
    ax.set_title(period, fontsize=6)
    ax.set_axis_off()
    markers = [plt.Line2D([0,0],[0,0],color=color, marker=".", linestyle="") for color in ["darkblue", "red"]]
    legend_labels = ["dis manibus (n={0})".format(str(len(df_edh_dm))), 
                     "dis manibus sacrum (n={0})".format(str(len(df_edh_dms)))]
    
    ax.legend(markers, legend_labels, numpoints=1, bbox_to_anchor=(0.6, 1), loc='upper left', fontsize=5)
   
plt.tight_layout(pad=0)
plt.subplots_adjust(wspace=0.0, hspace=0.0)
fig.suptitle(f'Spatial distribution of "dis manibus" formula by historic period (EDH dataset)', fontsize=8, fontweight="bold")
plt.savefig('../figures/04_EDH_Formulae_compared_production_periods.png')
fig
```



# Creating three snaphots in time

From simulated dates take 1-100, 101-200, 201-300 for future use.

```{r}
EDH_dated[80]
```

```{r}
EDH_1CAD<- EDH_dated %>% 
  select(id, longitude, latitude, type_of_inscription_clean, height_cm, width_cm, depth_cm, material_clean, type_of_monument_clean, province_label_clean, country_clean, findspot_ancient_clean, clean_text_conservative, clean_text_interpretive_word, clean_text_interpretive_word_lowercase, form_dis_manibus, random_dates, date_var_1) %>% 
  filter(date_var_1 > 0 & date_var_1 < 101) 
EDH_2CAD<- EDH_dated %>% 
  select(id, longitude, latitude, type_of_inscription_clean, height_cm, width_cm, depth_cm, material_clean, type_of_monument_clean, province_label_clean, country_clean, findspot_ancient_clean, clean_text_conservative, clean_text_interpretive_word, clean_text_interpretive_word_lowercase, form_dis_manibus, random_dates, date_var_1) %>%
  filter(date_var_1 > 100 & date_var_1 < 201)
EDH_3CAD<- EDH_dated %>% 
  select(id, longitude, latitude, type_of_inscription_clean, height_cm, width_cm, depth_cm, material_clean, type_of_monument_clean, province_label_clean, country_clean, findspot_ancient_clean, clean_text_conservative, clean_text_interpretive_word, clean_text_interpretive_word_lowercase, form_dis_manibus, random_dates, date_var_1) %>%
  filter(date_var_1 > 200 & date_var_1 < 301)
```

```{r}
# save the time sliced data locally
library(jsonlite)
EDH_1CADjson <- jsonlite::toJSON(EDH_1CAD, auto_unbox = TRUE, force = TRUE)
write(EDH_1CADjson, file="../data/EDH_DM_1CAD.json")
EDH_2CADjson <- jsonlite::toJSON(EDH_2CAD, auto_unbox = TRUE)
write(EDH_2CADjson, file="../data/EDH_DM_2CAD.json")
EDH_3CADjson <- jsonlite::toJSON(EDH_3CAD, auto_unbox = TRUE)
write(EDH_3CADjson, file="../data/EDH_DM_3CAD.json")
```




# Clustering of formulae in three time slices

```{r, 1CAD}
# Damerau-Levenshtein method - allows for substitution, deletion, insertion and transposition of individual characters
dl_distance_1CAD<- stringdistmatrix(EDH_1CAD$form_dis_manibus, EDH_1CAD$form_dis_manibus, method= "dl")
table(dl_distance_1CAD)
dl_distance_1CAD
# renaming rows and columns
idnames <- unique(as.character(EDH_1CAD$id))
rownames(dl_distance_1CAD) <- idnames
colnames(dl_distance_1CAD) <- idnames
# Clustering the distance results in a dendrogram
hc_1CAD <- hclust(as.dist(dl_distance_1CAD), method = "complete")
plot(hc_1CAD, main ="Clustering of 'Dis Manibus' inscriptions in the 1st c. AD", xlab="Clusters")
rect.hclust(hc_1CAD, k=3)
```

```{r, 2CAD}
# Damerau-Levenshtein method - allows for substitution, deletion, insertion and transposition of individual characters
dl_distance_2CAD<- stringdistmatrix(EDH_2CAD$form_dis_manibus, EDH_2CAD$form_dis_manibus, method= "dl")
table(dl_distance_2CAD)
dl_distance_2CAD
# renaming rows and columns
idnames <- unique(as.character(EDH_2CAD$id))
rownames(dl_distance_2CAD) <- idnames
colnames(dl_distance_2CAD) <- idnames
# Clustering the distance results in a dendrogram
hc_2CAD <- hclust(as.dist(dl_distance_2CAD), method = "complete")
plot(hc_2CAD, main ="Clustering of 'Dis Manibus' inscriptions in the 2nd c. AD", xlab="Clusters")
rect.hclust(hc_2CAD, k=3)
```

```{r, 3CAD}
# Damerau-Levenshtein method - allows for substitution, deletion, insertion and transposition of individual characters
dl_distance_3CAD<- stringdistmatrix(EDH_3CAD$form_dis_manibus, EDH_3CAD$form_dis_manibus, method= "dl")
table(dl_distance_3CAD)
dl_distance_3CAD
# renaming rows and columns
idnames <- unique(as.character(EDH_3CAD$id))
rownames(dl_distance_3CAD) <- idnames
colnames(dl_distance_3CAD) <- idnames
# Clustering the distance results in a dendrogram
hc_3CAD <- hclust(as.dist(dl_distance_3CAD), method = "complete")
plot(hc_3CAD, main ="Clustering of 'Dis Manibus' inscriptions in the 3rd c. AD", xlab="Clusters")
rect.hclust(hc_3CAD, k=3)
```


## Three networks in time

https://yulab-smu.top/treedata-book/chapter4.html
```{r}
# convert 'hclust' to 'phylo' object
phylo_tree <- as.phylo(hc_1CAD)
# get edges
graph_edges <- phylo_tree$edge
# get graph from edge list
graph_net <- graph.edgelist(graph_edges)
# plot graph
plot(graph_net)
```


```{r}
dl_distance_1CADdf<- as.data.frame(dl_distance_1CAD)
dl_distance_2CADdf<- as.data.frame(dl_distance_2CAD)
dl_distance_3CADdf<- as.data.frame(dl_distance_3CAD)
```

```{r, fig.width=30}
dl_distance_1CADdf[1:3,1:3]
library(igraph)
#network1 <- graph_from_incidence_matrix(dl_distance_1CADdf, weighted=TRUE)
#plot(network1)
```
```{r}
#network2 <- graph_from_incidence_matrix(dl_distance_2CADdf, weighted=TRUE)
#plot(network2)
```

```{r}
#network3 <- graph_from_incidence_matrix(dl_distance_3CADdf, weighted=TRUE)
#plot(network3)
```


## HeatMap in time


```{r}
#Load in temporal data from sciencedata

EDH_1CAD <-jsonlite::fromJSON("https://sciencedata.dk/public/b6b6afdb969d378b70929e86e58ad975/formulae/EDH_DM_1CAD.json")
EDH_2CAD <-jsonlite::fromJSON("https://sciencedata.dk/public/b6b6afdb969d378b70929e86e58ad975/formulae/EDH_DM_2CAD.json")
EDH_3CAD <-jsonlite::fromJSON("https://sciencedata.dk/public/b6b6afdb969d378b70929e86e58ad975/formulae/EDH_DM_3CAD.json")
```

```{r}
#Load in temporal data locally
#EDH_1CAD <- fromJSON("../data/EDH_DM_1CAD.json")
#EDH_2CAD <- fromJSON("../data/EDH_DM_2CAD.json")
#EDH_3CAD <- fromJSON("../data/EDH_DM_3CAD.json")
```

```{r}
library(leaflet.extras)
library(RColorBrewer)
heat_formulae1 <- EDH_1CAD %>% 
  leaflet(width="100%") %>%
  addTiles() %>% 
  addProviderTiles("Stamen.TerrainBackground")%>%
  addProviderTiles("Stamen.TonerBackground") %>% 
  setView( lng = 12.9239625, lat = 41.9515694, zoom = 4 ) %>%
  #setMaxBounds(lat1=40.633977, lng1 =-4.227926 , lat2=35.133882 , lng2=40.882336) %>%
 
  addHeatmap(lng = ~as.numeric(na.omit(EDH_1CAD$longitude)), lat = ~as.numeric(na.omit(EDH_1CAD$latitude)), 
            intensity = 0.5, layerId = NULL, group = NULL, minOpacity = 0.5, #max = 1,
  radius = 3, blur = 2, gradient = "Reds", cellSize = 1,) %>% 
  
 
  
  addLegend(position = "topright",
  colors = c("Red", "Blue", "Green"),
  labels = c("1st c. AD", "2nd c. AD", "3rd c. AD"), opacity = 1,
  title = "Heat map of Dis Manibus formulae" 
) %>% 
  addScaleBar(position="bottomleft")
heat_formulae1
```

```{r}
heat_formulae2 <- EDH_1CAD %>% 
  leaflet(width="100%") %>%
  addTiles() %>% 
  addProviderTiles("Stamen.TerrainBackground")%>%
  addProviderTiles("Stamen.TonerBackground") %>% 
  setView( lng = 12.9239625, lat = 41.9515694, zoom = 4 ) %>%
  #setMaxBounds(lat1=40.633977, lng1 =-4.227926 , lat2=35.133882 , lng2=40.882336) %>%
 
  addHeatmap(lng = ~as.numeric(na.omit(EDH_2CAD$longitude)), lat = ~as.numeric(na.omit(EDH_2CAD$latitude)), 
            intensity = 0.5, layerId = NULL, group = NULL, minOpacity = 0.5, #max = 1,
  radius = 3, blur = 2, gradient = "Blues", cellSize = 1,
  ) %>% 
  addScaleBar(position="bottomleft")
heat_formulae2
```
```{r}
heat_formulae3 <- EDH_3CAD %>% 
  leaflet(width="100%") %>%
  addTiles() %>% 
  addProviderTiles("Stamen.TerrainBackground")%>%
  addProviderTiles("Stamen.TonerBackground") %>% 
  setView( lng = 12.9239625, lat = 41.9515694, zoom = 4 ) %>%
  #setMaxBounds(lat1=40.633977, lng1 =-4.227926 , lat2=35.133882 , lng2=40.882336) %>%
 
  addHeatmap(lng = ~as.numeric(na.omit(EDH_3CAD$longitude)), lat = ~as.numeric(na.omit(EDH_3CAD$latitude)), 
            intensity = 0.5, layerId = NULL, group = NULL, minOpacity = 0.5, #max = 1,
  radius = 3, blur = 2, gradient = "Greens", cellSize = 1,
  ) %>% 
  addScaleBar(position="bottomleft")
heat_formulae3
```

