---
title: "Exploring Space and Time with the Hawkes process"
author: "Adela Sobotkova"
date: "16/08/2021"
output: html_document
---
## Initial setup

## Setup of the environment:

```{r setup, echo=TRUE, message=FALSE, warning = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(jsonlite)
```

# Guidelines for use
This script begins with the subsetting of the EDH_dis_manibus.json (11670x77) file down to the sample of inscriptions containing the Dis Manibus formula found in Spain. This is for easier temporal and spatial modelling. These filtering steps are recorded for reproducibility (and marked eval=false to prevent accidental reruns.

If interested in the spatio-temporal analysis, skip to line where you can read in the available EDH_DM_Spain csv. 

# Prepare the Data
## Loading data from Sciencedata.dk or local data folder 
Only use this if you need the original raw dataset (it is a sizable file ~300mb)
```{r, eval = false}
list_json <- jsonlite::fromJSON("https://sciencedata.dk/public/b6b6afdb969d378b70929e86e58ad975/EDH_dis_manibus_2021-08-12.json")
list_json <- jsonlite::fromJSON("../data/EDH_dis_manibus_2021-08-12.json")
EDH_DM <- as_tibble(list_json)
```

## Spanish subset of DIS MANIBUS records
```{r, eval = false}
# The original dataset is 11670 x 77 large, we need a simpler one
dim(EDH_DM)

# Cut columns to the essential ones
EDH_DMf <- EDH_DM %>% 
  select(id, not_before, not_after, longitude,latitude, people, type_of_monument,country, text_edition, type_of_monument_clean, type_of_inscription_clean,  country_clean,clean_text_interpretive_word, clean_text_interpretive_word_lowercase,clean_text_interpretive_sentence, form_dis_manibus)

DM_Spain <- EDH_DMf %>% 
  filter(country_clean == "Spain") %>% 
  select(-people) %>% 
  filter(!is.na(longitude))# lists are always an issue in spatial viz, so eliminating for now

write_csv(DM_Spain, "../data/EDH_DM_Spain.csv")
```

## Playing with time 
Hawkes usually expects time or date (YYYY-MM-DD) for calculations, so this is an exploration of 'what might work'?
```{r}
# Load the smaller sample dataset of DIS MANIBUS inscriptions from Spain
DM_Spain <- read_csv("../data/EDH_DM_Spain.csv")

# check the coordinates are numeric and non-missing
class(DM_Spain$latitude)
class(DM_Spain$longitude)

# does it look like Spain?
ggplot(DM_Spain, aes(x = longitude, y = latitude)) +
  geom_point()

# lots of messy dates in not_before column, which denotes the earliest date of creation
unique(DM_Spain$not_before)
       
# if you need yyyy/mm/dd, use this script (but you lose BC dates)       
DM_Spain <- DM_Spain %>% 
  filter(not_before != "-0030") %>% 
  filter(not_before != "-0050") %>% 
  filter(!is.na(not_before))
DM_Spain$dates <- paste0("1/1/",DM_Spain$not_before)

# maybe years will suffice if made numeric
DM_Spain <- DM_Spain %>% 
  filter(not_before != "-0030") %>% 
  filter(not_before != "-0050") %>% 
  filter(!is.na(not_before))
DM_Spain$t <- as.numeric(DM_Spain$not_before)
unique(DM_Spain$t)

# Make the 4 versions of Dis Manibus formula into a mark so we can check if these different versions impact one another or have different diffusion trajectories

unique(DM_Spain$form_dis_manibus)
DM_Spain$m <- factor(DM_Spain$form_dis_manibus, levels = c("dis manibus sacrum","dis manibus","d m", "d m s"))
levels(DM_Spain$m) <- c(1,2,3,4)
class(DM_Spain$m)
```

## Evaluating different factors in diffusion

Different factors may have influenced the dissemination of particular formulae throughout the Roman Empire, from elite competition to the location of stonemason workshops. In order to evaluate the role of different factors, we need to reflect these in the `marks` argument in Hawkes process (and most spatial analysis functions).  

`mark` denotes an attribute that contains a factor, or categorized values. 

In the case of DM inscriptions, I wish to create marks with which we could test for two alternative (exclusive) hypotheses about the diffusion of particular DM formulae:

1. Elite competition. Importance and wealth of the family/deceased individual manifest in the conspicuous nature of an inscription impacts the proliferation of the particular epigraphic expression (Arthur Saxe hypothesis). Size and quality of inscribed medium, length of inscription, unabridged content increase the overall expense of the production. The higher the expense, the more stimulus for local elite competition, and the greater proliferation of similar inscriptions in a desire to trump neighbours in advertising exploits. This approach should show high variability in elite production (long and expensive but differentiated inscriptions) and mimicry in non-elite production (freedmen and clients imitating the masters to the letters, but perhaps in more abridged, lower quality formats)

2. Workshops and regional fashion are the primary factors in the selection of particular formulae, resulting in (dialect-like) zones of use. We should see geographically constrained clusters where similarity decreases with distance from a *workshop*. The assumption here is that workshops do not have overlapping zones of influence??

There may be other options, of course, such as child workshops or masons working freelance, and creating variations on the canonical form...

In order to test these two hypotheses, I create several different marks for `expense` which encodes the cost of an inscription, based on the a)length of the inscription, b)dimensions of the medium, and c)the number of capitalised words within, d) osobni jmena z atributu people. 

### Length of inscription
One of the best ways to estimating the cost of an inscription is using its length, with number of words being a reasonable proxy. This calculation varies depending on whether it is run on transcription, or clean_text_interpretive_word, often inflating the former due to editorial notation. 

Let's compare the difference between transcription and interpretive word on this small example
```{r length-text-inspect}
# Check the editorial apparatus in the fourth inscription
head(EDH_tibble$transcription, 4)
# Compare the cleaned version
head(EDH_tibble$clean_text_interpretive_word)
# Count the words in fourth inscription
str_count(EDH_tibble$transcription[4], '\\w+')
# Count them in the clean version
str_count(EDH_tibble$clean_text_interpretive_word[4], '\\w+')

```
While there is some ambiguity given the number of capital letters, for the purpose of this analysis (lenght of text), we are happy with the more conservative estimate.

```{r length-text}
#Let's create a new column with the word counts.
EDH_tiny$wc_interpretive <- str_count(EDH_tibble$clean_text_interpretive_word, '\\w+')

# Inspect the result
summary(EDH_tiny$wc_interpretive) # few massive inscriptions
table(EDH_tiny$wc_interpretive) # there are a few zeroes - no inscriptions

# Which inscriptions have more than 1000 words?
which(EDH_tiny$wc_interpretive>1000)

# How many words do they contain?
EDH_tiny$wc_interpretive[EDH_tiny$wc_interpretive>1000]

# Let's display a histogram of wordcount log instead of wordcount to gauge the "magnitude"
hist(log10(EDH_tiny$wc_interpretive), main = "Logarithm of number of words on inscriptions")

# Write the log of length into the dataset
log10(0)

zeroes <- which(EDH_tiny$wc_interpretive == 0)
EDH_tiny$wc_interpretive[zeroes] <- NA
EDH_tiny$wc_interpretive_log <- log10(EDH_tiny$wc_interpretive)
summary(EDH_tiny$wc_interpretive_log)
```

### Size of the inscription medium
Second best way to estimating the cost of an inscription is to take into account the size of the stone block it is on (most epitaphs are on stone)

```{r size-inscription}
summary(EDH_tibble$height_cm)
summary(EDH_tibble$width_cm)
summary(EDH_tibble$depth_cm) # 10000 more inscriptions have NAs as in no depth

EDH_tibble %>% 
  select(height_cm,width_cm,depth_cm) %>% 
  head()

# Area
area_cm2<- EDH_tibble$height_cm*EDH_tibble$width_cm
length(which(is.na(area_cm2))) #36600 NAs
summary(area_cm2)

hist(log10(area_cm2))

# Volume
volume_cm3 <- EDH_tibble$height_cm*EDH_tibble$width_cm*EDH_tibble$depth_cm
summary(volume_cm3*0.001) # cubic decimeters, 46980 NAs 

# See the distribution
hist(log10(volume_cm3))

# Write to data
EDH_tiny$volume_cm3log <- log10(volume_cm3)
EDH_tiny$area_cm2log <- log10(area_cm2)
```

### Number of capitalized words
Another way of gauging the significance of an inscription is to count all the personal names. Important persons might have more names than middling merchants, or care to write them all out (e.g. emperors) engaging in conspicuous self-advertising, which we are trying to quantify here.

```{r}
# Compare the cleaned version: How many capitalized words do you see in the fourth inscription?
EDH_tibble$clean_text_interpretive_word[4]

# How many does R see?
str_count(EDH_tibble$clean_text_interpretive_word[4], '\\b[A-Z][a-z]+')


#Check the tail of the inscriptions
tail(EDH_tibble$clean_text_interpretive_word)

# What is actually on the stone?
tail(EDH_tibble$transcription)


# So, we seem to be erring on two oppositde sides: garbled or unexpanded abbreviations don't count, AND where god invocations appear, they amplify count.   Who knows what else happens. It is not a super good measure.

EDH_tiny$capitalized_words <- str_count(EDH_tibble$clean_text_interpretive_word, '\\b[A-Z][a-z]+')

EDH_tiny$capitalized_words_log <- log10(str_count(EDH_tibble$clean_text_interpretive_word, '\\b[A-Z][a-z]+')) 

hist(log10(EDH_tiny$capitalized_words), xlab = "number of capitalized words in inscription (log)", main = "Frequency of capitalized words in inscriptions (log)")
?hist()

```

# Join marks attributes with Spanish subset
```{r join-data}
# source datasets
dim(DM_Spain) # contains temporal and spatial data
dim(EDH_tiny) # contains attributes for 'marks'

# both datasets contain the shared key "id", which we can join by
DMS <- DM_Spain %>% 
  left_join(EDH_tiny, by = "id")
dim(DMS)

hist(DMS$wc_interpretive, main = "Length of inscriptions", xlab = "Number of words in an inscription")
```


# Analysis I
## Trying to use nphawkes package from P Boyde
credits: https://www.youtube.com/watch?v=lFhAQp8LHg4 

Peter is available now : fall 2021

What is the Hawkes process? It is a so-called self-exciting process in which the occurrence of a point temporarily elevates surrounding points. It is useful for cascading chains of events such as epidemics, social media memes, or stockexchange devaluations.  The assumption is that random background events spawn/trigger non-random children events in their (spatial?) vicinity.
 
In terms of probability: A Hawkes process (Hawkes, 1971; Liniger, 2009) supposes that past events can temporarily raise the probability of future events, assuming that such excitation is (1) positive, (2) additive over the past events, and (3) exponentially decaying with time.

```{r}
# install.packages("devtools")
# devtools::install_github("boydpe/nphawkes")
library(nphawkes)
?misd()
```

```{r}
# Trial run according to the documentation
data("hm.csv")
out = misd(dates = hm$t,
   ref_date = "1999-10-16",
   lat = hm$lat,
   lon = hm$lon,
   marks = hm$m,
   time_breaks = c(0,0.1, 0.5, 1,7,93,600), # days or seconds
   space_breaks = c(0,0.5, 1, 10, 25, 100), # entered as miles or kms
   mark_breaks = c(3, 3.1,3.3, 4, 5, 8),
   just_times = T)
out
View(out$p0) # see the probability (of observation being a parent of the variable) matrix
plot(out$time_breaks) # only your inputs
plot(out$h)

install.packages("cowplot")
library(cowplot)
trig_plots(out) # series of plots, h(s) is effect size
trig_plots(out, time_xlim = c(0, 15)) # reduce xlim to focus on effect - the ability of producing children over time in days/ kms in space / number of letter (in case of inscriptions)

```
g and h are probability density
k

space can be dropped


perc_diag 0.721/ 72.1% are probably parents or background events
amount of probability that exists along the diagonal

which is probabilistically a parent event?
Look at the mainshock and parent under "data" are calculated with the misd()

residual diagnostics: 
superthinning (included in the package) ; 
if the model is fitted well, in areas of high intensity , it should remove points; in areas of low intensity, it will simulate random points; in the end you get a uniform process.  If the output is not a uniform process, the thinning will not work correctly. 


## Run nphawkes::misd function on DIS MANIBUS dataset from Spain

```{r misd-length-inscription}
outDM <- misd(dates = DMS$t,
            lat = DMS$latitude,
            lon = DMS$longitude,
            marks = DMS$wc_interpretive_log,
            time_breaks = c(0,0.1, 0.25, 0.5,1,2.5, 5), # years
   space_breaks = c(0,10, 25, 50, 100, 250, 500))

# Let's look at inscription frequency through time
plot(table(DM_Spain$not_before))

# Now what does the Hawkes model actually show? 
hist(outDM$data$parent)
plot(outDM$g) # time triggering value for temporal bin; g and h are probability densities (in case of no bin assignment, a uniform split is applied). In the first bin, the event is more likely to spawn a child than in the next bins, until 6th point in time, than the probability drops. 
plot(outDM$h) # y axis is effect size; x axis should show what the effect is for each distance; play with the spatial binning; spread out the first bin;
View(outDM$p0)

trig_plots(outDM) 

```

### Area / Size of medium

```{r misd-medium-area}
# does not work because of missing values in marks/volume or area
outDM_size_cm2 <- misd(dates = DMS$t,
            lat = DMS$latitude,
            lon = DMS$longitude,
            marks = DMS$area_cm2log, #over 100 missing
            time_breaks = c(0,0.1, 0.25, 0.5,1,2.5, 5), # years
   space_breaks = c(0,10, 25, 50, 100, 250, 500))

summary(DMS$area_cm2log) #148 NAs
summary(DMS$volume_cm3log) #213 missing

trig_plots(outDM) 
```

## Capitalized word cound
```{r capit-word-count}
# does not work because of missing values in marks/volume or area
outDM_capitalized <- misd(dates = DMS$t,
            lat = DMS$latitude,
            lon = DMS$longitude,
            marks = DMS$capitalized_words, 
            time_breaks = c(0,0.1, 0.25, 0.5,1,2.5, 5), # years
   space_breaks = c(0,10, 25, 50, 100, 250, 500))

summary(DMS$capitalized_words) #213 missing

trig_plots(outDM_capitalized) 
```
## Questions for nphawkes

How do I know which are the probably parents and which are the children events?

What is the effect of the marks? (in our case the marks denote the different spellings of DIS MANIBUS which can reflect wealth and prestige of the commissioner)

Are the spatial locations taken into account? How?



# Analysis II
## Explore Hawkes package from Peter Halpin
credits: http://htmlpreview.github.io/?https://github.com/peterhalpin/hawkes/blob/master/hawkes_eg.html

```{r}
devtools::install_github("peterhalpin/hawkes")
library(hawkes)
head(email)
pp_obj <- pp(email)
print(pp_obj)

plot(pp_obj) # does not work (list of events by employer/employee)
gof(pp_obj)
```

### Simulate
```{r}

hawkes_em <- EM(pp_obj,nstarts = 1, maxit = 3)
hawkes_em

parms <- get_parms(hawkes_em)
gof(pp_obj, parms)

```

... To Be Continued

# Inspirational (visual) ppts and papers divulged by Google upon a quick search..
https://insurancedatascience.org/downloads/Zurich2019/Session4/3_Boumezoued_Alexandre.pdf
https://www.degruyter.com/document/doi/10.1515/phys-2020-0002/html
https://export.arxiv.org/pdf/2003.03671