---
title: "`Dis manibus`in time"
author: 
- Petra Hermankova^[Aarhus University, Denmark, https://orcid.org/0000-0002-6349-0540]
date: "`r format(Sys.Date())`"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged
---

```{r, echo=FALSE}
library(tidyverse)
library(jsonlite)
library(leaflet)
library(ggthemes)
library(rnaturalearth)
library(rnaturalearthdata)
library(tidytext)
library(tm)
library(quanteda)
library(topicmodels)
library(Matrix)
```


# Setup

**Loading data from Sciencedata.dk**

```{r}
#list_json <- jsonlite::fromJSON("https://sciencedata.dk/public/b6b6afdb969d378b70929e86e58ad975/EDH_text_cleaned_2021-01-21.json")
#EDH <- as_tibble(list_json)
```

**Loading data locally**
```{r}
# this will work only if you have the dataset available locally
EDH <- jsonlite::fromJSON("../data/EDH_text_cleaned_2021-01-21.json")
#dir.create("../figures")
```

## Text preparation

Converting the text of inscriptions to lowercase
```{r}
EDH <- EDH %>% 
  mutate(clean_text_interpretive_word_lowercase = str_to_lower(EDH$clean_text_interpretive_word)) %>% 
  mutate(clean_text_conservative_lowercase = str_to_lower(EDH$clean_text_conservative))
```

## Preparation coordinates for mapping
```{r}
# preparation of coordinates
EDH<- EDH %>% 
  separate(col = coordinates, into = c("longitude", "latitude"), sep = ",")

EDH$latitude <- as.numeric(str_replace(EDH$latitude, pattern = "\\)", replacement=""))
EDH$longitude <- as.numeric(str_replace(EDH$longitude, pattern = "c\\(", replacement=""))
```


# Epigraphic formulae on funerary inscriptions

Inspired by > Bennett and Laurence 2020: The Local and the Global: the Use of Formulae in the Epitaphs at Ostia

The following formulae typically occurs on funerary inscriptions.

Canonic Versions:

*Invocation to deities*
D M - Dis Manibus (invocation of the underworld ancestral spirits/deities)
D M S - Dis Manibus Sacrum

*Age aspect*
V A - Vixit annos (formula stating how many years the deceased lived)
P M - Plus minus (formula relating to uncertain age: more or less)

*Physical aspect of resting place*
H S E - Hic situs/sita est (formula introducing the deceased: here lies X)
IN AG IN FR - In Fronte ... In agros (formula specifying the extent of the burial plot)
H M H N S - Hoc Monumentum Heredem Non Sequetur (formula prohibiting future use of the burial plot by heirs)

*Personal quality of deceased and wishes*
B M - Bene merenti (formula praising the deceased)
L L P Q E - Libertis libertabusque posterisque eorum (formula for the freedmen and freedwomen and their descendants)
S T T L - Sit tibi terra levis (formula wishing the deceased: May the earth lie lightly on you)
O T B Q - Ossa tibi bene quiescant (formula wishing the deceased: may your bones rest well)

## Extracting inscriptions with for all *typical* variants of the formulae

```{r}
dms_regex <- "\\bdi{1,2}s manibus sacrum\\b"

dm_regex <- "\\bdi{1,2}s manibus\\b"

vixitannos_regex <- "\\bvixit ann[o|i]s\\b"

sittibi_regex <- "\\bsit tibi terra levis\\b"

hst_regex <- "\\bhic sita est|\\bhic situs est\\b|\\bhis siti sunt"

benemer_regex <- "\\bbene merenti\\b"

ifig_regex <- "\\bin fronte\\b"

llpe_regex <- "\\blibertis libertabus"

hmhns_regex <- "\\bheredem non sequetur\\b"

pm_regex <- "\\bplus minus\\b"

otbq_regex <- "\\bossa tibi bene quiescant\\b"

```

Extracting the formulae specified by the regular expressions into individual comlumns.
```{r}

EDH$form_dm <- ifelse(grepl(dm_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, dm_regex), NA)

EDH$form_dms <- ifelse(grepl(dms_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, dms_regex), NA)

EDH$form_va <- ifelse(grepl(vixitannos_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, vixitannos_regex), NA)

EDH$form_sttl <- ifelse(grepl(sittibi_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, sittibi_regex), NA)

EDH$form_hst <- ifelse(grepl(hst_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, hst_regex), NA)

EDH$form_bm <- ifelse(grepl(benemer_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, benemer_regex), NA)

EDH$form_ifig <- ifelse(grepl(ifig_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, ifig_regex), NA)

EDH$form_llps <- ifelse(grepl(llpe_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, llpe_regex), NA)

EDH$form_hmhns <- ifelse(grepl(hmhns_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, hmhns_regex), NA)

EDH$form_pm <- ifelse(grepl(pm_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, pm_regex), NA)

EDH$form_obtq <- ifelse(grepl(otbq_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, otbq_regex), NA)

```

Merging the formulas into one column by pivot longer function.
```{r}

#EDH %>% 
#  select(-formula_type, -name)

EDH<- EDH %>%
pivot_longer(
cols = starts_with("form_"),
values_to = "formula_type",
values_drop_na = TRUE
)

#Fixing hic situs est and hic sita est into one

EDH$formula_type <- ifelse(grepl("hic situs est|hic sita est", EDH$formula_type, ignore.case = T),  "hic sita/us est", EDH$formula_type)

#Fixing dis manibus and diis manibus est into one 

EDH$formula_type <- ifelse(grepl("diis|dis manibus\\z", EDH$formula_type, ignore.case = T),  "dis manibus", EDH$formula_type)

EDH$formula_type <- ifelse(grepl("dis manibus sacrum|diis manibus sacrum", EDH$formula_type, ignore.case = T),  "dis manibus sacrum", EDH$formula_type)

EDH$formula_type <- ifelse(grepl("vixit annis|annos", EDH$formula_type, ignore.case = T),  "vixit anno/is", EDH$formula_type)

EDH %>% 
  count(formula_type, sort=T)

```

## Subsetting only to funerary inscriptions
```{r}
#names(EDH)

EDHfun <- EDH %>% 
  select(id, formula_type, name, type_of_inscription_clean, type_of_monument_clean, material_clean, not_before, not_after, latitude, longitude, province_label_clean, clean_text_interpretive_word, clean_text_interpretive_word_lowercase, people) %>% 
  filter(type_of_inscription_clean == "epitaph")

write_rds(EDHfun, "../data/EDH_funerary.rds")
```

## Exploring formulae

```{r}
EDHfun %>% 
  count(formula_type, sort=T) %>% 
  mutate(ratio = round(n/(sum(n)/100),2))
```

### Province

#### Type of formula by province - frequence
```{r, fig.height=10, fig.width=10}
EDMfun_freq<- EDHfun %>% 
  #dplyr::filter(!is.na(formula_cat)) %>% 
  ggplot(aes(y=province_label_clean)) + 
  geom_bar(aes(fill=formula_type)) +
  #coord_cartesian(xlim=c(0,1700)) +
  theme_minimal() +
  theme(text = element_text(size=12)) +
  labs(y="Roman province", x="Formula", title= "Frequence of funerary formulae by province", subtitle = ggtitle(paste("n =", nrow(EDHfun), "funerary inscriptions (EDH)")))  
EDMfun_freq
ggsave("../figures/EDMfun_province_freq.png")

```

#### Ratio of formulae variants per province

NA - represents all inscriptions without any formula
```{r, fig.height=10}

form_form<- EDHfun %>% 
  select(formula_type, province_label_clean) %>% 
  count(formula_type, province_label_clean, sort=T)


EDHfun %>% 
  ggplot(aes(y=province_label_clean, fill=formula_type)) +
  geom_bar(position="fill") +
  theme_minimal() +
  theme(text = element_text(size=12)) +
  geom_vline(xintercept=c(0.25, 0.5, 0.75), linetype="dotted", color = "blue", size=0.8) + 
  labs(y="Roman province", x="Formula", title= "Ratio of funerary formulae by province", subtitle = ggtitle(paste("n =", nrow(EDHfun), "inscriptions (EDH)"))) 
  #geom_label(aes(label= form_dis_manibus)) +
  #geom_label(aes(label = form_dis_manibus), colour = "black", fontface = "bold", hjust = -0.1)
```

### Map of funerary formulae occurences - cummulative
```{r}
# preparation of subsets
# dis manibus formula
EDH_dm <- EDHfun %>% 
  filter(formula_type == "dis manibus")

# dis manibus sacrum formula
EDH_dms <- EDHfun %>% 
  filter(formula_type == "dis manibus sacrum")

# bene merenti formula
EDH_bm <- EDHfun%>% 
  filter(formula_type == "bene merenti")

# vixit annos formula
EDH_va <- EDHfun %>% 
  filter(formula_type == "vixit anno/is")

# plus minus formula
EDH_pm <- EDHfun %>% 
  filter(formula_type == "plus minus")

# hic sita/us est formula
EDH_hst <- EDHfun %>% 
  filter(formula_type == "hic sita/us est")

# libertis libertabus formula
EDH_llpe <- EDHfun %>% 
  filter(formula_type == "libertis libertabus")

# ossa tibi bene quiescant formula
EDH_otbq <- EDHfun %>% 
  filter(formula_type == "ossa tibi bene quiescant")

# sit tibi terra levis formula
EDH_stti <- EDHfun %>% 
  filter(formula_type == "sit tibi terra levis")

# heredem non sequetur formula
EDH_hmhns <- EDHfun %>% 
  filter(formula_type == "heredem non sequetur")

# in fronte formula
EDH_ifig <- EDHfun %>% 
  filter(formula_type == "in fronte")

```


```{r}
Formulae_map<- leaflet(EDHfun) %>% 
  leaflet(width="100%") %>%
  #addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
  addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
  #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
  #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  setView( lng = 12.9239625, lat = 41.9515694, zoom = 3.5 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
  #addPolylines(data = roads, color = "purple", weight = 1, opacity = 0.7) %>% 
  
  addCircles(lng = EDH_dm$longitude, 
             lat = EDH_dm$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "blue" , fillColor = "blue",) %>%
  addCircles(lng = EDH_dms$longitude, 
             lat = EDH_dms$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "red" , fillColor = "red",) %>%
  addCircles(lng = EDH_bm$longitude, 
             lat = EDH_bm$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "black" , fillColor = "brown",) %>%
  addCircles(lng = EDH_va$longitude, 
             lat = EDH_va$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "orange" , fillColor = "cyan",) %>%
  addCircles(lng = EDH_pm$longitude, 
             lat = EDH_pm$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "black" , fillColor = "yellow",) %>%
  addCircles(lng = EDH_stti$longitude, 
             lat = EDH_stti$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "orange" , fillColor = "green",) %>%
  addCircles(lng = EDH_hst$longitude, 
             lat = EDH_hst$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "black" , fillColor = "violet",) %>%
  addCircles(lng = EDH_ifig$longitude, 
             lat = EDH_ifig$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "orange" , fillColor = "pink",) %>%
  addCircles(lng = EDH_hmhns$longitude, 
             lat = EDH_hmhns$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "orange" , fillColor = "orange",) %>%
  addCircles(lng = EDH_llpe$longitude, 
             lat = EDH_llpe$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "orange" , fillColor = "white",) %>%
  addCircles(lng = EDH_otbq$longitude, 
             lat = EDH_otbq$latitude, opacity = 0.3, radius = 0.7, fill = T , color = "orange" , fillColor = "grey",) %>%
  addLegend(position = "bottomright",
  colors = c("Blue", "Green", "Red", "Orange"),
  labels = c("dis manibus", "diis manibus", "dis manibus sacrum", "diis manibus sacrum"), opacity = 1,
  title = "Dis Manibus formula (EDH)" 
) %>% 
  addScaleBar(position="bottomleft")

Formulae_map
```

```{r, fig.height=12}
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) +
    geom_sf() +
    geom_point(data = EDHfun, aes(x = longitude, y = latitude, color=formula_type), size = 1.5, 
        shape = 16, alpha = 0.5) +
    coord_sf(xlim = c(-10, 36), ylim = c(58,30), expand = TRUE) +
  facet_wrap(~formula_type, ncol=4) +
  labs(title = "Spatial distribution of common formulae on funerary inscriptions (synchronic)", subtitle = ggtitle(paste("n =", nrow(EDHfun), "inscriptions (EDH)")))  +
  theme_linedraw(base_size = 14)+
  theme(plot.title = element_text(size=40)) + 
  theme(axis.text = element_text(size = 18)) +
  theme(plot.subtitle = element_text(size=22)) +
  theme(strip.text.x = element_text(size =22)) +
  theme(legend.position = c(0.89, 0.15))



ggsave(filename = "../figures/EDH_formulas_map.png", width = 20, height = 16) 
```








# Temporal exploration of formulae

```{r, echo=FALSE}
#install.packages(c("reticulate", "stringdist", "ape", "igraph", "tidyverse", "reticulate", "jsonlite", "leaflet"))    

library(igraph)
library(ape)
library(stringdist)
library(tidyverse)
library(reticulate)
library(jsonlite)
library(leaflet)
#use_python("/usr/local/bin/python/") # use this if you have local python installed
#use_condaenv("pip")

Sys.which("python") # checks which python do I use
```
Sources: # http://amunategui.github.io/stringdist/

```{python}
# installation of packages is recommended to do via the terminal (see in your console below), pip install name-of-the-required-package

# https://vincent.doba.fr/posts/20210407_install-fiona-on-windows/ & https://geopandas.org/getting_started/install.html Issues with geopandas on windows

#to install requirements type and run in the terminal:  pip install -r requirements.txt

import fiona
import geopandas as gpd 
import requests
import seaborn as sns
import matplotlib.colors as mcolors
import matplotlib.pyplot as plt
from scipy.stats import trapz
import scipy
#import nltk
import json
import tempun
#import sddk
import numpy as np
import pandas as pd
pd.options.display.max_columns = 1000  # to see all columns
```


```{r}
# preparation of the dataset
EDHfun$not_before <- as.numeric(as.character(EDHfun$not_before))
EDHfun$not_after <- as.numeric(as.character(EDHfun$not_after))
```

```{python}
# How many inscriptions have both dates?
EDH_dated = r.EDHfun[(r.EDHfun["not_before"].notnull()) | (r.EDHfun["not_after"].notnull())]
len(EDH_dated)
```

```{python}
# Generate a list of 1000 random dates for each inscription in the dataset
EDH_dated["random_dates"] = EDH_dated.apply(lambda row: tempun.model_date(
    row["not_before"], row["not_after"], size=1000,), axis=1)
```

```{r}
# saving as r-object
py$EDH_dated -> EDH_dated
```

```{python}
# Count random dates in temporal timeblocks in a specified time range (from 200 BC to 600 AD, temporal blocks by 25 years)
EDH_complete_simulations = tempun.timeblocks_from_randoms(
    EDH_dated, "random_dates", [-200, 600, 25])
```

## Formulae and their types in time
```{python}
# Inscriptions by their formula type over time, EDH
simulations_by_type_len_EDH = []
for ins_type in r.EDHfun["formula_type"].unique():
    if ins_type != "NULL":
        subset_df = EDH_dated[EDH_dated["formula_type"] == ins_type]
        simulations_by_type_len_EDH.append((ins_type, len(subset_df)))
simulations_by_type_len_EDH = sorted(
    simulations_by_type_len_EDH, key=lambda x: x[1], reverse=True)
simulations_by_type_len_EDH
simulations_by_type_EDH = []
for ins_type_tup in simulations_by_type_len_EDH[:11]:
    subset_df = EDH_dated[EDH_dated["formula_type"]
                          == ins_type_tup[0]]
    simulations = tempun.timeblocks_from_randoms(
        subset_df, "random_dates", [-100, 600, 25])
    ins_type_N = len(subset_df)
    simulations_by_type_EDH.append([ins_type_tup[0], ins_type_N, simulations])
simulations_by_type_EDH = sorted(
    simulations_by_type_EDH, key=lambda x: x[1], reverse=True)
date_vars_by_instypes = []
for ins_type_tup in simulations_by_type_len_EDH[:11]:
    subset_df = EDH_dated[EDH_dated["formula_type"]
                          == ins_type_tup[0]]
    date_vars = []
    for n in range(100):
        date_vars.append(
            [date_var[n] for date_var in subset_df["random_dates"] if date_var != None])
    date_vars_by_instypes.append(date_vars)
```

```{python}
simulations_by_type_len_EDH
```

```{python}
# Figure 2
tab_colors_list = list(mcolors.TABLEAU_COLORS.keys())
plt.style.use("seaborn-white")
plt.rcParams['figure.dpi'] = 300
plt.rcParams['font.size'] = 10
fig, ax = plt.subplots(figsize=(14, 6), tight_layout=True)
for ins_type_sims, color in zip(simulations_by_type_EDH, tab_colors_list[:len(simulations_by_type_EDH)]):
    tempun.plot_timeblocks_data(ins_type_sims[2], ax=ax, color=color)
  
    
ax.set_xlabel("Year", weight="bold")
ax.set_ylabel("Count", weight="bold")
ax.set_title(f"'Formula types on funerary inscriptions over time (EDH, n={len(EDH_dated)})", weight="bold")
markers = [plt.Line2D([0, 0], [0, 0], color=color, lw=4)
           for color in tab_colors_list[:len(simulations_by_type_EDH)]]
legend_labels_EDH = [tup[0] + " (n={})".format(str(tup[1])) for tup in simulations_by_type_EDH]
ax.legend(markers, legend_labels_EDH, numpoints=1,  
           title=f"Formula types (EDH n={len(EDH_dated)})", title_fontsize="large", bbox_to_anchor=(1, 1), loc='upper right')
plt.savefig('../figures/02_EDH_Typologies_comparison_time.png')
fig
```




```{python}
def get_date_var(randoms):
    try:
        return randoms[0]
    except:
        return None
# selecting one random date out of the 1000 version
EDH_dated["date_var_1"] = EDH_dated["random_dates"].apply(get_date_var)
```

```{r}
# updating the values in existing r-object
py$EDH_dated -> EDH_dated
```

```{r}
EDH_1CAD<- EDH_dated %>% 
  filter(date_var_1 > 0 & date_var_1 < 100) 
EDH_2CAD<- EDH_dated %>% 
  filter(date_var_1 > 101 & date_var_1 < 200)
EDH_3CAD<- EDH_dated %>% 
  filter(date_var_1 > 201 & date_var_1 < 300)
EDH_4CAD<- EDH_dated %>% 
  filter(date_var_1 > 301 & date_var_1 < 400)

```

```{r}
# save the time sliced data locally
library(jsonlite)
EDH_1CADjson <- jsonlite::toJSON(EDH_1CAD, auto_unbox = TRUE)
write(EDH_1CADjson, file="../data/EDHfun_1CAD.json")
EDH_2CADjson <- jsonlite::toJSON(EDH_2CAD, auto_unbox = TRUE)
write(EDH_2CADjson, file="../data/EDHfun_2CAD.json")
EDH_3CADjson <- jsonlite::toJSON(EDH_3CAD, auto_unbox = TRUE)
write(EDH_3CADjson, file="../data/EDHfun_3CAD.json")
EDH_4CADjson <- jsonlite::toJSON(EDH_4CAD, auto_unbox = TRUE)
write(EDH_4CADjson, file="../data/EDHfun_4CAD.json")
```


## 1st century



```{r, fig.height=12}
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) +
    geom_sf() +
    geom_point(data = EDH_1CAD, aes(x = longitude, y = latitude, color=formula_type), size = 1.5, 
        shape = 16, alpha = 0.5) +
    coord_sf(xlim = c(-10, 36), ylim = c(58,30), expand = TRUE) +
  facet_wrap(~formula_type, ncol=4) +
  labs(title = "Spatial distribution of common formulae on funerary inscriptions (AD 1- 100)", subtitle = ggtitle(paste("n =", nrow(EDH_1CAD), "inscriptions (EDH), based on Monte Carlo simulation")))  +
  theme_linedraw(base_size = 14)+
  theme(plot.title = element_text(size=38)) + 
  theme(axis.text = element_text(size = 18)) +
  theme(plot.subtitle = element_text(size=22)) +
  theme(strip.text.x = element_text(size =22)) +
  #theme(legend.position = c(0.89, 0.15))
  theme(legend.position = "none")



ggsave(filename = "../figures/EDH_formulas_map_1CAD.png", width = 20, height = 16) 
```

## 2nd century

```{r, fig.height=12}
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) +
    geom_sf() +
    geom_point(data = EDH_2CAD, aes(x = longitude, y = latitude, color=formula_type), size = 1.5, 
        shape = 16, alpha = 0.5) +
    coord_sf(xlim = c(-10, 36), ylim = c(58,30), expand = TRUE) +
  facet_wrap(~formula_type, ncol=4) +
  labs(title = "Spatial distribution of common formulae on funerary inscriptions (AD 101-200)", subtitle = ggtitle(paste("n =", nrow(EDH_2CAD), "inscriptions (EDH), based on Monte Carlo simulation")))  +
  theme_linedraw(base_size = 14)+
  theme(plot.title = element_text(size=38)) + 
  theme(axis.text = element_text(size = 18)) +
  theme(plot.subtitle = element_text(size=22)) +
  theme(strip.text.x = element_text(size =22)) +
  #theme(legend.position = c(0.89, 0.15))
  theme(legend.position = "none")



ggsave(filename = "../figures/EDH_formulas_map_2CAD.png", width = 20, height = 16) 
```
## 3rd century

```{r, fig.height=12}
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) +
    geom_sf() +
    geom_point(data = EDH_3CAD, aes(x = longitude, y = latitude, color=formula_type), size = 1.5, 
        shape = 16, alpha = 0.5) +
    coord_sf(xlim = c(-10, 36), ylim = c(58,30), expand = TRUE) +
  facet_wrap(~formula_type, ncol=4) +
  labs(title = "Spatial distribution of common formulae on funerary inscriptions (AD 201-300)", subtitle = ggtitle(paste("n =", nrow(EDH_3CAD), "inscriptions (EDH), based on Monte Carlo simulation")))  +
  theme_linedraw(base_size = 14)+
  theme(plot.title = element_text(size=38)) + 
  theme(axis.text = element_text(size = 18)) +
  theme(plot.subtitle = element_text(size=22)) +
  theme(strip.text.x = element_text(size =22)) +
  #theme(legend.position = c(0.89, 0.15))
  theme(legend.position = "none")



ggsave(filename = "../figures/EDH_formulas_map_3CAD.png", width = 20, height = 16) 
```

## 4th century

```{r, fig.height=12}
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) +
    geom_sf() +
    geom_point(data = EDH_4CAD, aes(x = longitude, y = latitude, color=formula_type), size = 1.5, 
        shape = 16, alpha = 0.5) +
    coord_sf(xlim = c(-10, 36), ylim = c(58,30), expand = TRUE) +
  facet_wrap(~formula_type, ncol=4) +
  labs(title = "Spatial distribution of common formulae on funerary inscriptions (AD 301-400)", subtitle = ggtitle(paste("n =", nrow(EDH_4CAD), "inscriptions (EDH), based on Monte Carlo simulation")))  +
  theme_linedraw(base_size = 14)+
  theme(plot.title = element_text(size=38)) + 
  theme(axis.text = element_text(size = 18)) +
  theme(plot.subtitle = element_text(size=22)) +
  theme(strip.text.x = element_text(size =22)) +
  #theme(legend.position = c(0.89, 0.15))
  theme(legend.position = "none")



ggsave(filename = "../figures/EDH_formulas_map_4CAD.png", width = 20, height = 16) 
```

















# Textual similarity of funerary inscriptions

## Tokenizing words, splitting on an empty space
```{r}
EDH_tokenized <- EDHfun %>% 
  unnest_tokens(word, clean_text_interpretive_word, token = stringr::str_split, pattern = " ") %>% 
  drop_na(word) %>%
  print()
```

```{r}
#document term matrix

EDH_dtm <- EDH_tokenized %>%
  count(id, word) %>%
  cast_dtm(id, word, n)
EDH_dtm
```

```{r}
#document frequency matrix

EDH_dfm <- EDH_tokenized %>%
  count(id, word) %>%
  cast_dfm(id, word, n)
EDH_dfm
```

```{r}
#sparse matrix

EDH_sparse<- EDH_tokenized %>%
  count(id, word) %>% 
  cast_sparse(id, word, n)
dim(EDH_sparse)
```



## Exploration of matrix
```{r}
# how many words has epitaph
max(EDH_dfm["dis manibus",])
max(EDH_dfm["dis manibus sacrum",])
```


# Topic modelling

Latent Dirichlet allocation (LDA) is a particularly popular method for fitting a topic model. It treats each document as a mixture of topics, and each topic as a mixture of words. This allows documents to “overlap” each other in terms of content, rather than being separated into discrete groups, in a way that mirrors typical use of natural language.
```{r}
# topic modelling for 10 topics

EDH_lda_10 <- LDA(EDH_dtm, k=10, control=list(seed=1234))
```

```{r}
EDH_topics <- tidy(EDH_lda_10, matrix="beta")
```

```{r, fig.width=10}
EDH_top_terms <- EDH_topics %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>% 
  ungroup() %>%
  arrange(topic, -beta)
EDH_top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free_y", ncol=5) +
  scale_y_reordered()
```

Per type of formula classification. Each of these values is an estimated proportion of words from that document that are generated from that topic.

```{r}
EDH_dtm_formula <- EDH_tokenized %>%
  count(formula_type, word) %>%
  cast_dtm(formula_type, word, n)

EDH_lda_5_formula <- LDA(EDH_dtm_formula, k=5, control=list(seed=1234))

EDH_gamma <- tidy(EDH_lda_5_formula, matrix = "gamma")

EDH_gamma %>% 
  mutate(title = reorder(document, gamma * topic)) %>%
  ggplot(aes(factor(topic), gamma)) +
  geom_boxplot() +
  facet_wrap(~ document) +
  labs(x = "topic", y = expression(gamma))
# continue here https://www.tidytextmining.com/topicmodeling.html
```


## Similarity matrix

document-term matrix. Having the data in this form, we can calculate distances between  the  vectors  by  measuring  and  inverting  their  cosine  similarity.  This  way  we  obtain  a  matrix  expressing distance between any two works within the corpus, with works sharing a larger proportion of words being less remote to each other than works employing less overlapping vocabulary. This distance matrix could be finally projected into a 2-dimensional space by using t-distributed Stochastic Neighbor Embedding (tSNE) and plotted as a scatter plot (see Fig. 1 below).

https://github.com/sdam-au/PIA/blob/master/scripts/3_OVERVIEW%2BWORK-DISTANCES.ipynb




# Experiments with extraction of additional formulae

```{r}
rip_regex <- "\\brequiesc[i|e|a|an]t in pace" #more strict


rip_regex <- "\\brequies.* [in pace]" # more open

EDH$form_rip <- ifelse(grepl(rip_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, rip_regex), NA)
length(EDH$form_rip)
#table(EDH$form_rip)
```

```{r}
def_regex <- "\\bdefunct[a|us|i] \\w+" #more strict


EDH$form_def <- ifelse(grepl(def_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, def_regex), NA)
table(EDH$form_def)
```
```{r}
testam_regex <- "\\bex testament." #more strict


EDH$form_testam <- ifelse(grepl(testam_regex, EDH$clean_text_interpretive_word_lowercase, ignore.case = T),  str_extract(EDH$clean_text_interpretive_word_lowercase, testam_regex), NA)
table(EDH$form_testam)
```